---
title: "Penguins_Zoe"
author: "Zoe Zilz"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in penguins.csv
```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read.csv("penguins.csv")
```
###Our question is: Is there a significant difference in mean flipper lengths for female chinstrap, gentoo, and adelie penguins near Palmer Archipelago

How many factors do I have? 
1 - Species
3 levels - chinstrap, gentoo, adelie
1 continuous random variable - flipper length

More than 2 groups for continuous measured data -- maybe I can use a one way ANOVA?
assume: independent observations - sure
assume: equal variance - will use Levene's test for
assume: normality - use histograms and qq plots to explore this

First, want to know how many observations are in each group (why?)
can use table function to calculate counts based on however many factors you give it

```{r}

View(penguins)
counts <- with(penguins, table(species,sex))

#updating dataset because Allison doesn't like penguin names
# using case when to add a column called "common name"

penguins2 <- penguins %>% 
  mutate(
    common_name =
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "chinstrap",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "adelie",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "gentoo"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex =="FEMALE")
```
Exploratory graphs and Levene's test for equal variances

```{r}
# maybe grouped differently?
hist1 <- ggplot(penguins2, aes( x = flipper_length))+
  geom_histogram(aes(fill = common_name))

hist1
#yeah!

#normally distributed?
hist2 <- ggplot(penguins2, aes( x = flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~common_name, scale = "free")

hist2


#looks pretty norm

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~common_name, scale = "free")

qqs

#based on qq plots and hists, obs for each group appear normall distributed. also n > 30 for all groups, so means will be normally distributed by CLT
```

