---
title: "Penguins_Zoe"
author: "Zoe Zilz"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in penguins.csv
```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read.csv("penguins.csv")
```
###Our question is: Is there a significant difference in mean flipper lengths for female chinstrap, gentoo, and adelie penguins near Palmer Archipelago

How many factors do I have? 
1 - Species
3 levels - chinstrap, gentoo, adelie
1 continuous random variable - flipper length

More than 2 groups for continuous measured data -- maybe I can use a one way ANOVA?
assume: independent observations - sure
assume: equal variance - will use Levene's test for
assume: normality - use histograms and qq plots to explore this

First, want to know how many observations are in each group (why?)
can use table function to calculate counts based on however many factors you give it

```{r}

counts <- with(penguins, table(species,sex))

#updating dataset because Allison doesn't like penguin names
# using case when to add a column called "common name"

penguins2 <- penguins %>% 
  mutate(
    common_name =
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "chinstrap",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "adelie",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "gentoo"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex =="FEMALE")
```
Exploratory graphs and Levene's test for equal variances

```{r}
# maybe grouped differently?
hist1 <- ggplot(penguins2, aes( x = flipper_length))+
  geom_histogram(aes(fill = common_name))

hist1
#yeah!

#normally distributed?
hist2 <- ggplot(penguins2, aes( x = flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~common_name, scale = "free")

hist2


#looks pretty norm

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~common_name, scale = "free")

qqs

#based on qq plots and hists, obs for each group appear normall distributed. also n > 30 for all groups, so means will be normally distributed by CLT
```

Evaluate variance and perform a Levene's Test for equal variances

If the largest sample variance is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

```{r}

#using group_by
variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances # the largest variance is definitely not more than 4x the smallest variance

# Levene's Test
# leveneTest function in 'car'

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin

# H0: no difference in variances (equal)
# HA: Variances are NOT equal

# Here, there is not significant differnece in variances across species p > 0.05
```

One-Way ANOVA for penguin flipper lengths by species (common_name)

```{r}
# ASKING R: please compare means based on the rand variable flipper_length as a function of common_name, and you're going to get that from penguins2

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# then full report using summary to get anova outcome

# H0: Means across all groups are equal
# HA: At least two group means differ significantly

#follow up, post hoc, which differ?

#Tukey's HSD (works nicely with aov() function )

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph
```
###Type of statement you can make
Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post hoc Tukey's HSD, $\alpha$ = 0.05)
